---
title: "Class Reference"
author: "Derek Willis"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

### How to set the working directory

The command to set a working directory is setwd(). For example, this sets the working directory to a directory called data_journalism_2024_fall inside a directory called jour472 located inside your home directory (the ~ is equivalent of /Users/[youruser] on a Mac).

```{r}
setwd("~/jour472/data_journalism_2024_fall")

getwd()

```

### How to install and load the tidyverse

```{r}
#install.packages("tidyverse")
library(tidyverse)
```

Operators

  |> continue
  | or
  & and
  <- assign
  = equals
  > - greater than
  >= - greater than or equal to
  %in% - equal to in cases of multiple variables
  ^ - starts with
  | - or

Functions
  Lab 1
  
  glimpse() - shows a glimpse of the columns in the data
  arrange() - arranges the data
    desc() - orders data in descending order

  Lab 2
  
  select() - allows you to show specific columns
  str_to_upper() - makes text uppercase
  case_when() - organizes into categories
  nrow/ncolumn - gives the number of rows/columns
  mdy() and its combinations - turns data into a date
  filter() - allowes you to filter data by row
  
  Lab 3
  
  guess_max() - uses the first 10 rows to set the data type
  problem() - checks for problems on loading data
  as.numeric - changes a columns data type to number
  clean_names() - standardizes column names
  reanme() - renames a column
  get_dupes() - checks for duplicate rows
  distinct() - gets rid of duplicate rows
  use str_tot_title() - standardizes capitalization
  floor_date() - sorts from lowest earliest date
  
  Lab 4 
  
  str_detect() - detects values in columns
  key_collision_merge - attempts to summarize similar values
  write_csv() = writes to a csv file
  str_remove - removes words from data
  str_detect - shows specific words
  
  Lab 5
  
  bind_rows() - combines rows from different dataframes
  list() - does something inside of bind_rows()
  left_join - takes all the records from the left table and only the records that match in the right one
  right_join - does the same thing
  inner_join - takes only the records where they are equal
  full_join - returns all rows of both, regardless of if there’s a match
  join_by() - joins by correct columns
  