---
title: "lab_10"
author: "Derek Willis"
date: "2024-11-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

-   the tidyverse, janitor and tigris libraries
-   you might choose to use the tidycensus library for Q3

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load libraries
library(tidyverse)
library(janitor)
library(sf)
library(tigris)
```

Let's explore the election results from last night and make some maps!

## Questions

**Q1.** Make a county-level map of Maryland's results for U.S. Senate using `md_senate_county_24.csv` in the data folder, calculating the difference between Angela Alsobrooks's percentage of the total votes and Larry Hogan's percentage of the total votes and displaying that percentage difference on the map. What is the story here, and what is the county with the most interesting results?

**A1.** Usually democrats are able to run away with votes in highly populated areas but Hogan was able to keep it close in places like Baltimore County and Anne Arundel.

```{r}
md_senate <- read_csv("data/md_senate_county_24.csv")

md_senate <- md_senate |>
  mutate(
    total = Alsobrooks + Hogan + Scott,
    per_Alsobrooks = (Alsobrooks/total)*100,
    per_Hogan = (Hogan/total)*100,
    per_diff = per_Alsobrooks - per_Hogan,
    GEOID = as.character(GEOID)
  )

md_counties <- counties() |>
  filter(STATEFP == "24")

md_vote_senate <- md_counties |>
  left_join(md_senate, join_by("GEOID")) |>
  clean_names()

ggplot() +
  geom_sf(data=md_vote_senate, aes(fill=per_diff)) +
  scale_colour_viridis_b(option="magma") +
  theme_minimal()
```

**Q2.** Make a county-level map showing the difference between Donald Trump's county-level performance this year and Larry Hogan's, using percentages to compare the two as you did in Q1. You can use the dataframe you initially loaded for Q1, and you'll need to load the 2024 presidential results in the data folder and join them to it before proceeding. Are there any counties where Trump got a higher percentage than Hogan? How would you describe the map showing the Trump-Hogan difference?

Also answer this: is a map the best way to present this data? What else could you make that might convey more information?

**A2.** Trump barely gets more votes from Allegany, but that's super negliable, everything else he is down five to ten percent-ish. I think a map is an alright way of conveying the information, but maybe a line graph would better show the difference between the two.

```{r}
md_pres <- read_csv("data/md_pres_county_24.csv") |>
  mutate(GEOID = as.character(GEOID)) |>
  mutate(
    total_nat = Harris + Trump + Oliver + Stein + Kennedy,
    per_Trump = (Trump/total_nat)*100,
  )

md_diff_nat <- md_pres |>
  left_join(md_senate, join_by("County", "GEOID")) |>
  mutate(
    per_diff_nat = per_Hogan - per_Trump
  )

md_vote_nat <- md_counties |>
  left_join(md_diff_nat, join_by("GEOID")) |>
  clean_names()

ggplot() +
  geom_sf(data=md_vote_nat, aes(fill=per_diff_nat)) +
  scale_colour_viridis_b(option="magma") +
  theme_minimal()
```

**Q3** Make another map showing the difference between Larry Hogan's county-level performance this year and from his governor race in 2018, using percentages to compare the two as you did in Q2. You can use the dataframe you initially loaded for Q1, and you'll need to load the 2018 governor results in the data folder and join them to it before proceeding. Are there any counties where Hogan did better this year? How would you describe the map showing the difference?

**A3** No, it appears that Hogan has done worse this year compared to 2018.

```{r}
library(mapview)
```

```{r}
md_gov <- read_csv("data/md_gov_county_18.csv") |>
  adorn_totals("col") |>
  as.data.frame() |>
  mutate(
    per_Hogan_18 = (Hogan/Total)*100,
    GEOID = as.character(GEOID)
    ) |>
  select(GEOID, per_Hogan_18)

md_vote_18v24 <- md_gov |>
  left_join(md_senate, join_by(GEOID)) |>
  mutate(Hogan_diff = per_Hogan - per_Hogan_18)

md_vote_18v24c <- md_counties |>
  left_join(md_vote_18v24, join_by("GEOID")) |>
  clean_names()

md_vote_18v24c |>
  ggplot() +
  geom_sf(aes(fill=hogan_diff)) +
  scale_colour_viridis_b(option="magma") +
  theme_minimal()
```

**Q4.** Choose your own map adventure! In the data folder I've included Maryland county-level results for the abortion rights amendment vote this year, plus the 2020 presidential results by county and 2022 governor's results by county. Using one of those and at least one other contest for comparison, make a county-level map showing what you think is the most interesting aspect of that data, and describe the results in terms of a possible story. You can bring in other data, such as Census information, if you like.

**A4.** I looked at the difference between Trump votes and the votes against abortion. The amount of people voting for Trump is significantly higher than the amount of people who voted against abortion. The difference becomes closer as the places become more heavily populated.

```{r}
md_abortion <- read_csv("data/md_abortion_24.csv") |>
  adorn_totals("col") |>
  mutate(
    per_against = (Against/Total)*100
  )

md_abort_pres <- md_pres |>
  left_join(md_abortion, join_by(County)) |>
  mutate(per_abort_diff = per_Trump - per_against)

md_abort_presc <- md_counties |>
  left_join(md_abort_pres, join_by(GEOID))

md_abort_presc |>
  ggplot() +
  geom_sf(aes(fill=per_abort_diff)) +
  scale_colour_viridis_b(option="magma") +
  theme_minimal()
```

-30-
